R Course Syllabus

### 2023-04-26 Draft Syllabus.Rmd
Begin writing lesson 6

### Topics Covered

@@

### Lesson

@@

1. What is a function?
2. When to use functions.
3. Writing functions in R.

```{r lesson6}

>>>>

### **Topics Covered**

@@

### **Lesson**

@@

### **Topics Covered**

1. What is a function?
    + When to use functions.
2. Writing functions in R.
3. When to write a function.

### **Lesson**

#### **Functions Defined**

You have been using functions since you began working in R. `paste()`, `mutate()`, `sum()`, and even `%>%` are examples of functions. Fundamentally, a function just contains some uniform code that can be applied multiple times without typing. Theoretically you do not need to write functions, but neglecting them make your code far longer, less readable, and less easy to understand. Functions are incredibly important for repetitive or complex work.

Let's look at some examples:

Adding numbers:
```{r l6_1_1}
# You can add numbers directly, using the Mathematical operators we discussed in Lesson 3
4 + 4

# But you can also just use the sum function
sum(4, 4)

# This does not save much time with 2 numbers, but what about with 25
some_numbers <- round(runif(25, 0, 100))

# With addition signs
1+31+55+78+24+78+46+65+73+74+86+3+74+31+65+51+55+45+41+21+86+44+54+62+71

# With sum
sum(some_numbers)
```

Creating new variables:
```{r l6_1_2}
# Built-in data set
mtcars

# Creating variables
mtcars$cyl_mpg <- mtcars$cyl/mtcars$mpg

# Creating variables with mutate
mtcars <- mtcars %>%
  mutate(cyl_mpg2 = cyl/mpg)

# Identical. Time savings can be seen if you want to create more than one variable
mtcars[, c("cyl_mpg", "cyl_mpg2")]

# Add 2 to every variable with mutate and across
mtcars %>%
  mutate(across(everything(), ~ . + 2))
```

Performing multiple data transformations:
```{r l6_1_3}
# Reset mtcars data set
rm(mtcars)

# Keep only mpg
mtcars <- data.frame(mpg = mtcars$mpg)

# Multiply mpg by 2
mtcars$mpg2 <- mtcars$mpg * 2

# Keep if mpg2 greater than or equal to 40
(mtcars <- mtcars[mtcars$mpg2 >= 40, ])

# Same steps, using the pipe
rm(mtcars)

mtcars %>%
  # Keep only mpg
  select(mpg) %>%
  # Multiply mpg by 2
  mutate(mpg2 = mpg*2) %>%
  # Keep if mpg2 greater than or equal to 40
  filter(mpg2 >= 40)
```

Functions make working in R a lot easier and often more intuitive. You should use them as often as makes sense. If you find yourself repeating tasks, you should consider replacing that code with a function. This is especially true if the task is complex, using a function guarantees that you are applying the same code every time you want to perform that task.

#### **Writing Funcitons**

You can create a function (often called it *defining* the function) using the following syntax:

name_of_function <- function(arguments_of_function) {
  operations performed by function
}

The first thing you will type will be the name that you want to give to your function. You can think of this as assigning the code that follows the assignment operator `<-` to the name. That name will now be associated with that code.

Next, you will write `function()`, and inside the parentheses you will write all of the arguments that can be provided to your function. For example, the function `is.na()` takes just one argument, x, which can be a list, atomic vector, or data frame (as well as some other object classes); the function `str_to_upper()`  has two arguments, a string and a locale.

Then you will open a curly brace (and when you type the first one in R studio, it will automatically create a second for you; press enter and it will also nicely split them across lines for you). On a new line, begin writing what you want your function to do. When you are done with the function definition, on a new line type a closing brace (this is likely unnecessary if you are using R studio, since it has probably already created the closing brace for you). Run the entire block of code, starting from the name through the ending curly brace and you will have created a function!

Here are some simple example functions:

### **Helpful Links**
```{r lesson6}
